\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-16 paper/paper 2/dc-2015.tex
na

\subsection{\emph{Simple Constraints}}

The \emph{minimum qualified cardinality restriction (R-75)} {\small\ms{Publication $\sqsubseteq$ $\geq$1 author.Person,}}
which restricts publications to have at least one author which must be a person,
is an example of a simple constraint on \emph{author}
which holds for all individuals of the class \emph{Publication}.
Table \ref{tab:property-constraint-cardinality-restriction} displays how the simple constraint is generically represented using the RDF-CV. Note that instead of the description logics construct $\geq$, the constraining element \emph{minimum qualified cardinality restriction} may be used.

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Minimum Qualified Cardinality Restriction as Property Constraint}
	\label{tab:property-constraint-cardinality-restriction}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left property list} & \textbf{right p. list} & \textbf{classes} & \textbf{constraining element} & \textbf{c. value} \\
      \hline
      Publication & author & - & Person & $\geq$ & 1 \\
		\end{tabular}
\end{table}

@@

For property constraints, \emph{left} and \emph{right property lists} are specified.
The assignment of properties to these lists happens relative to the constraining element.
\emph{Object Property Paths} (\emph{R-55})
ensure that if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the author of the book \emph{A-Brief-History-Of-Time} 
%{\small\ms{(authorOf(Stephen-Hawking, A-Brief-History-Of-Time))}} 
whose genre is \emph{Popular-Science}, 
%{\small\ms{(genre(A-Brief-History-Of-Time, Popular-Science)),}}
the object property path {\small\ms{authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre}} infers that \emph{Stephen-Hawking} is an author of the genre \emph{Popular-Science}. 
%{\small\ms{(authorOfGenre(Stephen-Hawking, Popular-Science))}.}
Thus, when representing the property constraint using the RDF-CV (see Table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are placed on the left side of the constraining element \emph{property paths}
and the property \emph{authorOfGenre} on its right side. As this property constraint holds for all individuals within the data, the context class is set to the \emph{DL top concept} $\top$ which stands for the super-class of all possible classes.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Object Property Paths as Property Constraint}
	\label{tab:property-constraint-object-property-paths}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      $\top$ & authorOf, genre & authorOfGenre & $\top$ & property path & - \\
		\end{tabular}
\end{table}

@@

Table \ref{tab:simple-constraint-not-expressible-in-dl)} presents (1) in the first line how the literal pattern matching simple constraint which is not expressible in DL and (2) in the second line how the universal quantification complex constraint which is expressible in DL are represented using the RDF-CV. Thereby, the context class \emph{ISBN}, whose instances must satisfy the simple constraint, is reused within the list of classes the complex constraint refers to. The literal pattern matching constraint type introduces the constraining element \emph{regex} whose validation has to be implemented once like for any other constraining element.
%In general, validation has to be implemented once for each \emph{generic constraint type} which is not expressible in DL. 
%\begin{gcotable}
%property & Enterprise & registryNumber & - & RegistryNumber & $\forall$ & - \\
%class & RegistryNumber & - & - & xsd:string & regex & 'NCC-1701([-][A-E])?' \\
%\end{gcotable}
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simple Constraints which are not Expressible in DL}
	\label{tab:simple-constraint-not-expressible-in-dl)}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      ISBN & - & - & xsd:string & REGEX & '\string^\text{$\backslash$d$\{9\}$[$\backslash$d$\mid$X]}\$' \\
      Book & identifier & - & ISBN & universal quantification & - \\
		\end{tabular}
\end{table} %\tb{Erman, can you please format the regular expression in the table above correctly?}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simplified Complex Constraints}
	\label{tab:simplified-complex-constraints}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
Publication & - & - & E, F & exclusive or & - \\
E & - & - & A, B & intersection & - \\
F & - & - & C, D & intersection & - \\
A & isbn & - & string & = & 1 \\
B & title & - & string & = & 1 \\
C & issn & - & string & = & 1 \\
D & title & - & string & = & 1 \\
		\end{tabular}
\end{table}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Primary Key Properties as Complex Constraints}
	\label{tab:primary-key-properties-as-complex-constraints}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
$\top$ & isbn$^{-}$ & isbn$^{-}$ & - & inverse property & - \\
Book & isbn$^{-}$ & - & - & functional property & - \\
		\end{tabular}
\end{table}

@@

\begin{ex}[commandchars=\\\{\}]
[   a \textit{rdfcv:SimpleConstraint} ;
    \textit{rdfcv:contextClass} Book ; 
    \textit{rdfcv:leftProperties} ( isbn ) ; 
    \textit{rdfcv:constrainingElement} "primary key" ] .
\end{ex}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Irreflexive Object Properties as Complex Constraints}
	\label{tab:irreflexive-object-properties-as-complex-constraints}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
$\exists$ authorOf.Self & authorOf & - & Self & existential quantification & - \\
$\neg$ $\exists$ authorOf.Self & - & - & $\exists$ authorOf.Self & negation & - \\
$\top$ & - & - & $\top$, $\neg$ $\exists$ authorOf.Self & sub-class & - \\
		\end{tabular}
\end{table}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Irreflexive Object Properties as Simple Constraints}
	\label{tab:irreflexive-object-properties-as-simple-constraints}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
$\top$ & authorOf & - & - & irreflexive property & - \\
		\end{tabular}
\end{table}

@@

\subsection{Bidirectional Constraint Transformations}

@@

\subsection{XXX}

\begin{ex}
owl:Thing 
    spin:rule [ a sp:Construct ; sp:text """
        CONSTRUCT {            
            ?cc
                a owl:Restriction ;
                owl:minQualifiedCardinality ?cv ;
                owl:onProperty ?lp1 ;
                owl:onClass ?c1 . }
        WHERE { 
            ?this
                a rdfcv:SimpleConstraint ;
                rdfcv:contextClass ?cc ;
                rdfcv:leftProperties ?leftProperties ;
                rdfcv:classes ?classes ;
                rdfcv:constrainingElement "minimum qualified cardinality restriction" ;
                rdfcv:constrainingValue ?cv .  
            ?leftProperties 
                rdf:first ?lp1 ;
                rdf:rest rdf:nil .    
            ?classes 
                rdf:first ?c1 ;
                rdf:rest rdf:nil . } """ ; ] .
\end{ex}

>>>>

\subsection{\emph{Simple Constraints}}
In this and the following section, we provide examples for the representation of constraint types within the framework.

The \emph{minimum qualified cardinality restriction (R-75)} {\small\ms{Publication $\sqsubseteq$ $\geq$1 author.Person,}}
which restricts publications to have at least one author which must be a person,
is an example of a simple property constraint on \emph{author}
which holds for all individuals of the class \emph{Publication}.
Table \ref{tab:property-constraint-cardinality-restriction} displays how the property constraint is generically represented using the RDF-CV.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Minimum Qualified Cardinality Restriction as Property Constraint}
	\label{tab:property-constraint-cardinality-restriction}
	\centering
		\begin{tabular}{c|c|c|c|c|c|c}
      \textbf{constraint set} & \textbf{context class} & \textbf{left property list} & \textbf{right p. list} & \textbf{classes} & \textbf{constraining element} & \textbf{c. value} \\
      \hline
      property & Publication & author & - & Person & $\geq$ & 1 \\
		\end{tabular}
\end{table}

@@

For property constraints, \emph{left} and \emph{right property lists} are specified.
The assignment of properties to these lists happens relative to the constraining element.
\emph{Object Property Paths} (\emph{R-55})
ensure that if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the author of the book \emph{A-Brief-History-Of-Time} 
%{\small\ms{(authorOf(Stephen-Hawking, A-Brief-History-Of-Time))}} 
whose genre is \emph{Popular-Science}, 
%{\small\ms{(genre(A-Brief-History-Of-Time, Popular-Science)),}}
the object property path {\small\ms{authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre}} infers that \emph{Stephen-Hawking} is an author of the genre \emph{Popular-Science}. 
%{\small\ms{(authorOfGenre(Stephen-Hawking, Popular-Science))}.}
Thus, when representing the property constraint using the RDF-CV (see Table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are placed on the left side of the constraining element \emph{property paths}
and the property \emph{authorOfGenre} on its right side. As this property constraint holds for all individuals within the data, the context class is set to the \emph{DL top concept} $\top$ which stands for the super-class of all possible classes.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Object Property Paths as Property Constraint}
	\label{tab:property-constraint-object-property-paths}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      property & $\top$ & authorOf, genre & authorOfGenre & $\top$ & property paths & - \\
		\end{tabular}
\end{table}

@@

Table \ref{tab:simple-constraint-not-expressible-in-dl)} presents (1) in the first line how the literal pattern matching simple constraint which is not expressible in DL and (2) in the second line how the universal quantification complex constraint which is expressible in DL are represented using the RDF-CV. Thereby, the context class \emph{ISBN}, whose instances must satisfy the simple constraint, is reused within the list of classes the complex constraint refers to. The literal pattern matching constraint type introduces the constraining element \emph{regex} whose validation has to be implemented once like for any other constraining element.
%In general, validation has to be implemented once for each \emph{generic constraint type} which is not expressible in DL. 
%\begin{gcotable}
%property & Enterprise & registryNumber & - & RegistryNumber & $\forall$ & - \\
%class & RegistryNumber & - & - & xsd:string & regex & 'NCC-1701([-][A-E])?' \\
%\end{gcotable}
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simple Constraints which are not Expressible in DL}
	\label{tab:simple-constraint-not-expressible-in-dl)}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      class & ISBN & - & - & xsd:string & regex & '\string^\text{$\backslash$d$\{9\}$[$\backslash$d$\mid$X]}\$' \\
      property & Book & identifier & - & ISBN & universal quantification & - \\
		\end{tabular}
\end{table} %\tb{Erman, can you please format the regular expression in the table above correctly?}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Simplified Complex Constraints}
	\label{tab:simplified-complex-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
class & Publication & - & - & E, F & XOR of property groups & - \\
class & E & - & - & A, B & intersection & - \\
class & F & - & - & C, D & intersection & - \\
property & A & isbn & - & string & = & 1 \\
property & B & title & - & string & = & 1 \\
property & C & issn & - & string & = & 1 \\
property & D & title & - & string & = & 1 \\
		\end{tabular}
\end{table}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Primary Key Properties as Complex Constraints}
	\label{tab:primary-key-properties-as-complex-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. set} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
property & $\top$ & isbn$^{-}$ & isbn$^{-}$ & - & inverse property & - \\
property & Book & isbn$^{-}$ & - & - & functional property & - \\
		\end{tabular}
\end{table}

@@

\begin{ex}[commandchars=\\\{\}]
[   a \textit{rdfcv:PropertyConstraint} , \textit{rdfcv:SimpleConstraint} ;
    \textit{rdfcv:contextClass} Book ; \textit{rdfcv:leftProperties} ( isbn ) ; \textit{rdfcv:constrainingElement} "keyfor" ] .
\end{ex}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Irreflexive Object Properties as Complex Constraints}
	\label{tab:irreflexive-object-properties-as-complex-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
property & $\exists$ authorOf.Self & authorOf & - & Self & existential quantification & - \\
class & $\neg$ $\exists$ authorOf.Self & - & - & $\exists$ authorOf.Self & negation & - \\
class & $\top$ & - & - & $\top$, $\neg$ $\exists$ authorOf.Self & sub-class & - \\
		\end{tabular}
\end{table}

@@

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Irreflexive Object Properties as Simple Constraints}
	\label{tab:irreflexive-object-properties-as-simple-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
property & $\top$ & authorOf & - & - & irreflexive property & - \\
		\end{tabular}
\end{table}

@@

\subsection{Mapping implementation}

@@
